Design of Adaptive Constellations and Error Protection Coding for Wireless Network Coding in 5-node Butterfly Networks

### 2017-05-26 Const_design.tex
Completed section 7.4 and added relevant point to intro

In this paper we extend the communication scheme from \cite{Prochazka-Uricar-Halls-Sykora_2015-ICC}
by appending appropriate \emph{error protection coding} to the system,
resulting in a practical channel-coded adaptive communication scheme
for WBN. We provide the following results and contributions:
\begin{enumerate}
\item We summarize the \emph{systematic constellation design algorithm}
from \cite{Prochazka-Uricar-Halls-Sykora_2015-ICC} and we provide
additional numerical results analysing its promising performance.
\item We revise and extend the results discussing the \emph{adaptive capabilities
of the constellations, }including \emph{the constellation mapping
regions} for adaptive WBN relaying\emph{.}
\item We validate the proposed design in a \emph{real-world HW setup}, including
the HW evaluation of adaptive system performance.
\item We show how the \emph{channel coding} can be integrated into the proposed
constellation design, including the analysis of proper source transmission
rates.
\item We evaluate numerically the performance of the resulting \emph{adaptive
modulation and coding} scheme for a wide range of channel conditions
in the WBN.
\item We show that the proposed design is viable even if the real-world
conditions induce some deviation from the assumed system model.
\end{enumerate}
The rest of the paper is organized as follows. The system model, including
all necessary definitions and details about WNC processing in WBNs
is summarized in Section~\ref{sec:CTUpp_SystemModel}. A systematic
\emph{constellation design algorithm} for WBNs is revised in Section~\ref{sec:CTUpp_Euclidean-Distance-based-Constel}.
The performance analysis of the proposed constellations (including
the HW evaluation) in an uncoded system is available in Section~\ref{sec:CTUpp_Performance-Analysis-Uncoded}.
Section~\ref{sec:CTUpp_Adaptive-Constellation-Design} discusses
an adaptive constellation design and Section~\ref{sec:CTUpp_Channel-Coding}
introduces the channel coding extension to the system. The performance
of the novel adaptive modulation and coding scheme is analyzed in
Section~\ref{sec:CTUpp_Performance-Analysis-Coded}, including a
simple robustness analyzis. Concluding remarks are summarized in Section~\ref{sec:CTUpp_DiscussionConclusions-etc}.

@@

\subsection{Analysis of Transmission Rates}

@@

\subsection{Three-Step Reference Scenario}
For a more advanced comparison of the proposed method, we consider two conventional \emph{3-step}
communication scenarios.

The first scenario is equivalent to a conventional routing, where the sources subsequently transmit
their data packet towards the relay. Relay then sends a complete source information (no NC coding is
applied) and thus the HSI channels are completely ignored (see Figure \ref{fig:CTUpp_ref3stepRouting}).

In the second scenario both sources again transmit their information one by one, but the relay is
allowed to combine the received (fully decoded) data packets into the network coded packet
$f(d_{A},d_{B})=d_{A}\oplus d_{B}$ and HSI channels are necessary to retrieve the desired
information at both sources (see Figure \ref{fig:CTUpp_ref3stepNC}).

To provide a fair comparison, in both reference scenarios we assume that the sources transmit at
double power, as they transmit time is essentially halved when compared with the 2-step scenarios.

@@

We compared throughput of the proposed adaptive constellation design from Section
\ref{sub:Maximal-Throughput-of} on the USRP devices with optimized tuple ($r_b, r_s, N_b, N_s$) for
each measured point (HW-pre-rotation on) with the reference scenarios in Figures \ref{fig:CTUpp_comparison-schemaa} and
\ref{fig:CTUpp_comparison-schemab}. We also considered the proposed method with
pre-rotation turned off as in Section \ref{sub:Robustness_Analysis} for comparison. Finally, we
added two theoretical limits for the 3-step scenarios, where ideal Gaussian alphabets are assumed.

@@

One can observe that

>>>>

In this paper we extend the communication scheme from \cite{Prochazka-Uricar-Halls-Sykora_2015-ICC}
by appending appropriate \emph{error protection coding} to the system,
resulting in a practical channel-coded adaptive communication scheme
for WBN. We provide the following results and contributions:
\begin{enumerate}
\item We summarize the \emph{systematic constellation design algorithm}
from \cite{Prochazka-Uricar-Halls-Sykora_2015-ICC} and we provide
additional numerical results analysing its promising performance.
\item We revise and extend the results discussing the \emph{adaptive capabilities
of the constellations, }including \emph{the constellation mapping
regions} for adaptive WBN relaying\emph{.}
\item We validate the proposed design in a \emph{real-world HW setup}, including
the HW evaluation of adaptive system performance.
\item We show how the \emph{channel coding} can be integrated into the proposed
constellation design, including the analysis of proper source transmission
rates.
\item We evaluate numerically the performance of the resulting \emph{adaptive
modulation and coding} scheme for a wide range of channel conditions
in the WBN.
\item We show that the proposed design is viable even if the real-world
conditions induce some deviation from the assumed system model.
\item We compare the proposed design with state of the art reference scherios to highlight its
superior performance.
\end{enumerate}
The rest of the paper is organized as follows. The system model, including
all necessary definitions and details about WNC processing in WBNs
is summarized in Section~\ref{sec:CTUpp_SystemModel}. A systematic
\emph{constellation design algorithm} for WBNs is revised in Section~\ref{sec:CTUpp_Euclidean-Distance-based-Constel}.
The performance analysis of the proposed constellations (including
the HW evaluation) in an uncoded system is available in Section~\ref{sec:CTUpp_Performance-Analysis-Uncoded}.
Section~\ref{sec:CTUpp_Adaptive-Constellation-Design} discusses
an adaptive constellation design and Section~\ref{sec:CTUpp_Channel-Coding}
introduces the channel coding extension to the system. The performance
of the novel adaptive modulation and coding scheme is analyzed in
Section~\ref{sec:CTUpp_Performance-Analysis-Coded}, including a
simple robustness analyzis. Concluding remarks are summarized in Section~\ref{sec:CTUpp_DiscussionConclusions-etc}.

@@

\subsection{Analysis of Transmission Rates}\label{Sec:Transmission_Rates}

@@

\subsection{Benchmarking with the State of the Art Three-Step Scenarios}
For a more advanced comparison of the proposed method, we consider two conventional \emph{3-step}
communication scenarios.

\subsubsection{Reference Three-Step Scenarios}\label{Sec:Reference-3step-models}
The first scenario both sources again transmit their information one by one, but the relay is
allowed to combine the received (fully decoded) data packets into the network coded packet
$f(d_{A},d_{B})=d_{A}\oplus d_{B}$ and HSI channels are necessary to retrieve the desired
information at both sources, that is the traditional network coding approach (see Figure
\ref{fig:CTUpp_ref3stepNC}).

In the scenario is equivalent to a conventional routing, where the sources subsequently transmit
their data packet towards the relay. Relay then sends a complete source information (no NC coding is
applied) and thus the HSI channels are completely ignored (see Figure \ref{fig:CTUpp_ref3stepRouting}).

\subsubsection{Implementation Details for Reference Scenarios}
To provide a fair comparison, in both reference scenarios we assume that the sources transmit at
double power, as they transmit time is essentially halved when compared with the 2-step scenarios.
We assume QAM($N$) constellation and adaptively look for $N, r$ maximizing the throughput as in the
proposed Section \ref{sub:Maximal-Throughput-of}.

\subsubsection{Maximum Transmission Rates for the Three-Step Reference Scenarios}
The theoretical limits for the considered 3-step scenarios can be easily derived (see
\cite{Uricar-Qian-Sykora-Mow_2014-IEEE-T-WC}) as
\begin{eqnarray}
  r^{\mathrm{3step}}_{\mathrm{route}} &\le& \min\left\{\frac{2\mathrm{Ca}(\gamma_{MAC})}{2},\frac{2\mathrm{Ca}(\gamma_{BC})}{2}\right\}\label{eq:cap3strout}\\
  r^{\mathrm{3step}}_{\mathrm{NC}} &\le& \min\left\{\frac{2\mathrm{Ca}(2\gamma_{MAC})}{2},\frac{2\mathrm{Ca}(\gamma_{HSI})}{2}, \mathrm{Ca}(\gamma_{BC})\right\},\label{eq:cap3stNC}
\end{eqnarray}
where $\mathrm{CA}(\gamma)=\log_2(1+\gamma)$ denotes conventional point to point channel capacity for Gaussian
channel \cite{Cover-Thomas_ElemInfTheory}.

These bounds serve as an initial step for the tuple (rate, alphabet cardinality) optimization as in
Section \ref{Sec:Transmission_Rates} but also as an ultimate theoretical bound of the individual
reference strategies.

@@

\subsubsection{Numerical Results}
Within this experiment, we evaluated on our USRP platform the following situations:
\begin{itemize}
  \item The proposed adaptive constellation design from Section \ref{sub:Maximal-Throughput-of} 
  with optimized tuple ($r_b, r_s, N_b, N_s$) for each measured point
  (HW-pre-rotation on).
  \item The proposed adaptive constellation design with violated assumptions (pre-rotation off)
  \item Both 3-step reference scenarios introduced in Section \ref{Sec:Reference-3step-models}.
  \item Fundamental theoretical limits for the conventional reference 3-step scenarios.
\end{itemize}

We considered two setups:
\begin{itemize}
  \item Strong HSI and equal MA and BC. This scenario is suited for standard wireless network
  coding, because it can efficiently utilize the strong HSI channel. See Figure
  \ref{fig:CTUpp_comparison-schemaa} for the corresponding result. As one can see, the proposed
  framework significantly outperforms both reference 3-step scenarios as well the corresponding
  theoretical limits even if the pre-rotation is off! The bottleneck for the 3-step routing is BC
  channel, because it needs to pass twice amount of data from relay to destinations compared to
  all other schemes. The 3-step NC approach does not suffer from this and therefore its performance
  is a bit better, nevertheless we pay for orthogonality in the MA stage. The proposed performance
  of the proposed framework is superior to both reference models.
  \item Strong BC channel refers (among others) to a case with high relay power. The result is shown
  in Figure \ref{fig:CTUpp_comparison-schemab}. The 3-step routing works much better compared to the previous
  case, because of the strong BC channel. The conventional 3-step network coding suffers from weak
  HSI channel. The proposed design again significantly outperforms both reference 3-step scenarios.
  We should stress, that the performance gain is also caused by a \emph{partial usage} of the weak
  HSI channel.
\end{itemize}

@@
